Hazelnut: A Bidirectionally Typed \\ Structure Editor Calculus

### 2016-06-28 hazelnut-popl17.tex
adding plausible statements of construct and reach

Reachability\todo{state reachability theorem}

@@

Constructibility\todo{state constructibility theorem}
\end{subequations}

>>>>

\paragraph{Reachability}

We give the following theorem to demonstrate that the rules for movement do
indeed capture the process of moving the focus to any editable position
within a term. Intuitively, given two zippered expressions that are the
same up to the erasure of focus, its proof gives a list of composable
actions that, when applied to the first term produces the second.

As a consequence of applying Theorem \ref{thrm:actsafe} inductively at
every action in such a list, we couldn't hope to construct an ill-typed
term, so the statement here matches the mutually recursive structure of the
bidirectional typing judgements.

\begin{theorem}[Reachability]\todo{why $\htau$ and not $\ztau$?}
\label{thrm:reach}
Both of the following hold:
\begin{enumerate}
  \item If $\hsyn{\hGamma}{\removeSel{\zexp}}{\htau}$ and
   $\hsyn{\hGamma}{\removeSel{\zexp'}}{\htau}$ and $\removeSel{\zexp}
   = \removeSel{\zexp'}$ then $\exists L \in \mathit{list}
   ~\mathtt{action}$ such that $\mathit{iterate}~ L~ \zexp
   = \zexp'$.

  \item If $\hana{\hGamma}{\removeSel{\zexp}}{\htau}$ and
   $\hana{\hGamma}{\removeSel{\zexp'}}{\htau}$ and $\removeSel{\zexp}
   = \removeSel{\zexp'}$ then $\exists L \in \mathit{list}
   ~\mathtt{action}$ such that $\mathit{iterate}~ L~ \zexp
   = \zexp'$.
\end{enumerate}
\end{theorem}

@@

\paragraph{Constructibility}

We give the following theorem to demonstrate that the rules for the action
$\aConstruct{\varphi}$ do indeed capture the process of building any term
of the form $\varphi$. Intuitively, its proof gives a list of composable
actions that, when applied one sequentially to an empty hole in focus,
produce any well-typed expression.

As with Theorem \ref{thrm:reach}, the statement of this theorem must match
the mutually recursive structure of the bidirectional typing judgements.

\begin{theorem}[Constructability]\todo{why $\htau$ and not $\ztau$?}
\label{thrm:construct}
Both of the following hold:
\begin{enumerate}
  \item If $\hsyn{\hGamma}{\hexp}{\htau}$ then $\exists \zexp'$ $\exists
      L \in \mathit{list} ~\mathtt{action}$ such that
      $\mathit{iterate}~ L~ \zwsel{\hhole{}} = \zexp'$ and
      $\removeSel{\zexp'} = \hexp$

  \item If $\hana{\hGamma}{\hexp}{\htau}$ then $\exists \zexp'$ $\exists
      L \in \mathit{list} ~\mathtt{action}$ such that $\mathit{iterate}~
      L~ \zwsel{\hhole{}} = \zexp'$ and $\removeSel{\zexp'} = \hexp$
\end{enumerate}
\end{theorem}

Because actions are defined on the zippered form of expressions, we argue
that we can construct a term that is the same as the target up to the
erasure of focus rather than on the nose. Since any form inside a zippered
term may be in focus, there are linearly many such terms for a given
$\hexp$---Theorem \ref{thrm:construct} can be composed with
Theorem \ref{thrm:reach} to reach any one that might be desired in
particular.

\end{subequations}
